---
title: "Data Manipulation"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

This document will show how to *manipulate* data

Import the two datasets that we're going to manipulation 

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ",", "."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data_import_examples/FAS_pups.csv", na = c("NA", ",", "."))

pups_df = 
    janitor::clean_names(pups_df)

```

## `Select`

Use `select()` to select variables 

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

selecting everything from group to gd18_weight
```{r}
select(litters_df, group:gd18_weight)
```

can select to remove things that you don't want 
```{r}
select(litters_df, -pups_survive)
```

can combine
```{r}
select(litters_df, -(group:gd18_weight))
```

can select for suffixes
```{r}
select(litters_df, starts_with("gd"))

select(litters_df, contains ("pups"))
```

can rename as I go (select() selects out a group and renames it) rename() keeps the entire dataframe but renames the column
```{r}
select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)
```

these two do the same thing to relocate a column, just need to indicate that keeping everything in select
```{r}
select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight)
```


##`filter`



##`mutate`
