---
title: "Data Import"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

This document will show how to import data

#Import the FAS litters CSV 

Import the dataset and clean the names

```{r}
litters_df = read_csv("data_import_examples/FAS_litters.csv")

litters_df = janitor::clean_names(litters_df)
```

## Look at the dataset 

```{r}
litters_df

head(litters_df)
tail(litters_df, 10)
```

```{r eval = FALSE}
view(litters_df)

```

## Learnign assessment import with FAS_pups.csv absolute and relative paths

Use relative paths

```{r}
pups_df = read_csv("data_import_examples/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)

pups_df

```

Use absolute path

```{r, eval = FALSE}
pups_df = read_csv("/Users/florapang/Desktop/Data Science/20240917 Data Wrangling i/data_import_examples/FAS_pups.csv")
```

## Look at read_csv options 

Skip rows and turn off column names 

```{r}
litters_df = 
  read_csv(
    file = 'data_import_examples/FAS_litters.csv',
    col_names = FALSE,
    skip = 1
  )
```

## What about missing data

NA, empty cells and . are all supposed to be seen as missing values

```{r}
litters_df = 
  read_csv(
    file = 'data_import_examples/FAS_litters.csv',
    na = c("NA", "", ".")
  )

#calculate the mean of something but it hass missing values

litters_df = janitor::clean_names(litters_df)

pull(litters_df, gd0_weight)
```

What if we code 'group' as a factor variable? (current is a character)

```{r}
litters_df = 
  read_csv(
    file = "data_import_examples/FAS_litters.csv",
    na = c("NA", "", "."),
    col_types = cols(
      Group = col_factor()
    )
  )
```


## Import an excel file

Import MLB 2011 summary data

```{r}

#need to load another library other than tidyverse, best practice to load all 
#libraries in the same setup chunk

mlb_df = read_excel("data_import_examples/mlb11.xlsx", sheet = 'mlb11')
```

## Import DAD data

```{r}
pulse_df = read_sas("data_import_examples/public_pulse_data.sas7bdat")
```

## Never use read.csv()

Use read_csv (tibble) rather than read.csv

```{r}
litters_df = read_csv("data_import_examples/FAS_litters.csv")

litters_df$S
```

