---
title: "Data Tidying"
output: github_document
---

Select and pull only pulling out numbers (more rare) where as select will select a column 

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```

```{r}
pulse_df = 
  read_sas("data_import_examples/public_pulse_data.sas7bdat")
  janitor::clean_names()
```

This needs to go from wide to long format
Here the columns are values (here data rectangle where the depression indicators BDI are all spread across 4 columns according to their visits, where as want all the BDI scores in one column)

```{r}
pulse_tidy_df =
  pulse_df %>% 
  pivot_longer(
    cols = BDIScore_BL:BDIScore_01m:BDIScore_06m:BDIScore_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "BDIScore_"
  )
```

The names_prefix takes off the BDIScore_ prefix so that the data becomes cleaner and not every entry has the prefix 

Or could do differently with mutate

```{r}
pulse_tidy_df =
  pulse_df %>% 
  pivot_longer(
    cols = BDIScore_BL:BDIScore_01m:BDIScore_06m:BDIScore_12m,
    names_to = "visit",
    values_to = "bdi_score",
    names_prefix = "BDIScore_"
  ) %>% 
  mutate(
    visit = replace(visit, visit == "BL", "00m")
  ) %>% 
  relocate(Id, visit)
```

Do one more example.
Here weight are spread across two columns both are gestational day 
Need to make there only one gestational weight column, (from 2 columns and 1 row to 1 column to 2 rows)

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ".", "")) %>% 
  janitor::clean_names()
```

```{r}
litters_tidy_df =
  litters_df %>% 
    janitor::clean_names() %>% 
  pivot_longer(
    cols = gd0_weight:gd18_weight,
    names_to = "gd_time",
    values_to = "weight"
  ) %>% 
  mutate(
    gd_time = case_match(
      gd_time,
      "gd0_weight" ~ 0,
      "gd18_weight" ~18
    ))
```

Case match, any time you see this value you want to rename it to something else
In gd_time, anytime seeing gd0_weight it becomes 0

## Pivot wider

Let's make up an analysis results table 

```{r}
analysis_df = 
  tibble(
    group = c("treatment", "treatment", "control", "control"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 10, 4.2, 5)
  )
```

Need to pivot wider for human readability (currently good for coding but hard to draw conclusions)

```{r}
analysis_df %>% 
  pivot_wider(
    names_from = time,
    values_from = mean
  ) %>% 
  knitr::kable()
```

