---
title: "Data Manipulation"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

This document will show how to *manipulate* data

Import the two datasets that we're going to manipulation 

```{r}
litters_df = 
  read_csv("data_import_examples/FAS_litters.csv", na = c("NA", ",", "."))

litters_df = janitor::clean_names(litters_df)

pups_df = 
  read_csv("data_import_examples/FAS_pups.csv", na = c("NA", ",", "."))

pups_df = 
    janitor::clean_names(pups_df)

```

## `Select`

Use `select()` to select variables 

```{r}
select(litters_df, group, litter_number, gd0_weight)
```

selecting everything from group to gd18_weight
```{r}
select(litters_df, group:gd18_weight)
```

can select to remove things that you don't want 
```{r}
select(litters_df, -pups_survive)
```

can combine
```{r}
select(litters_df, -(group:gd18_weight))
```

can select for suffixes
```{r}
select(litters_df, starts_with("gd"))

select(litters_df, contains ("pups"))
```

can rename as I go (select() selects out a group and renames it) rename() keeps the entire dataframe but renames the column
```{r}
select(litters_df, GROUP = group)

rename(litters_df, GROUP = group)
```

these two do the same thing to relocate a column, just need to indicate that keeping everything in select
```{r}
select(litters_df, litter_number, gd0_weight, everything())

relocate(litters_df, litter_number, gd0_weight)
```

Learning Assessment

```{r}
select(pups_df, litter_number, sex, pd_ears)
```


##`filter`

Select works on columns, filter() filters out rows
Taking a variable inside the dataset, asking if it is true or false, if it is true then keep if false then do not

```{r}
filter(litters_df, gd_of_birth == 20)
```
double equal == ask whether this equal to that, asks a question (ask whether something is true about a variable)

greater than, greater than equal to, not equal to (numeric variables)
```{r}
filter(litters_df, pups_born_alive > 8)
filter(litters_df, pups_born_alive >= 8)
filter(litters_df, pups_born_alive != 9)
```

looking at character variables
```{r}
filter(litters_df, group == "Low8")
filter(litters_df, group %in% c("Low7", "Low8"))
```
not using the == since asking whether it is in Low7 or Low8 (question is or)

```{r}
filter(litters_df, group %in% c("Low7", "Low8"), pups_born_alive == 8)
```
here, two things true at the same time (either Low7 or Low8) AND pups born alive equals to 8

This is case sensitive, need to be specific about lower and upper case

to get rid of rows with missing data
doing it on everything will drop all rows with missing variables
```{r}
drop_na(litters_df)
```
or could specify

```{r}
drop_na(litters_df, gd0_weight)
```

Notice that we are not saving any of the new data frame (we are always going back to the original litters_df)
normally could do new_litters_df = drop_na(litters_df, gd0_weight) to create a new data frame

Learning assessment 
```{r}
filter(pups_df, sex == 1)
filter(pups_df, pd_walk < 11, sex == 2)
```


##`mutate`


##`arrange`
